---
title: "individual2.Rmd"
author: "Andrew Fahmy"
date: "3/16/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(readxl)
library(tidyverse)
library(ggplot2)
defense <- read_excel('cyclonesFootball2020.xlsx', sheet='Defensive')
offense <- read_excel('cyclonesFootball2020.xlsx', sheet='Offensive')

```


```{r}
defense <- defense %>%
  mutate(Name = as.factor(Name), 
         Opponent_Opponent = as.factor(Opponent_Opponent))

offense <- offense %>%
  mutate(Name = as.factor(Name), 
         Opponent_Opponent = as.factor(Opponent_Opponent))

```

```{r}
defense <- defense %>%
  mutate(across(Tackles_Solo:Pass_PB, as.numeric))

offense <- offense %>%
  mutate(across(Receiving_REC:Passing_INT, as.numeric))

```


```{r}
brock = colSums(offense %>% filter(Name == "Purdy, Brock") %>% select(-c("Name", "Opponent_Opponent")), na.rm = TRUE)
team = colSums(offense %>% filter(Name != "Purdy, Brock") %>% select(-c("Name", "Opponent_Opponent")), na.rm = TRUE)

dat = data.frame(brock, team)
dat = data.frame(t(dat))
dat

datt = offense
datt$brock = datt$Name=="Purdy, Brock"

datt  %>% ggplot(aes(x=1,y=Passing_YDS, fill=brock)) + geom_col() + coord_polar(theta = 'y')
datt  %>% ggplot(aes(x=1,y=Passing_TD, fill=brock)) + geom_col() + coord_polar(theta = 'y')


```
As we can see from the data and the graph, almost all of the offensive yards
thrown as well as touch downs are from Brock. This shows that the teams
performance is extremely reliant on the performance of brock.


```{r}
defense1 <- read_excel('cyclonesFootball2019.xlsx', sheet='Defensive')
offense1 <- read_excel('cyclonesFootball2019.xlsx', sheet='Offensive')
biography1 <- read_excel('cyclonesFootball2019.xlsx', sheet='Biography')

defense1 <- defense1 %>%
  mutate(across(Tackles_Solo:Pass_PB, as.numeric))

offense1 <- offense1 %>%
  mutate(across(Rushing_ATT:Passing_INT, as.numeric))

defense1 <- defense1 %>%
  mutate(Name = as.factor(Name), 
         Opponent_Opponent = as.factor(Opponent_Opponent))

offense1 <- offense1 %>%
  mutate(Name = as.factor(Name), 
         Opponent_Opponent = as.factor(Opponent_Opponent))

offense2 = offense %>% filter(Name %in% offense1$Name)
offense3 = offense1 %>% filter(Name %in% offense$Name)

mean1 = aggregate(x=offense2$Receiving_YDS, by=list(offense2$Name), FUN=mean)
mean1$ave_2019 = aggregate(x=offense3$Receiving_YDS, by=list(offense3$Name), FUN=mean)$x

mean1$improvement = mean1$ave_2019 - mean1$x 

mean1[order(mean1$improvement), ]

defense2 = defense %>% filter(Name %in% defense1$Name)
defense3 = defense1 %>% filter(Name %in% defense$Name)

mean2 = aggregate(x=defense2$Tackles_Solo, by=list(defense2$Name), FUN=mean)
mean2$ave_2019 = aggregate(x=defense3$Tackles_Solo, by=list(defense3$Name), FUN=mean)$x

mean2$improvement = mean2$ave_2019 - mean2$x

mean2[order(mean2$improvement), ]

```
We can see that for offense, Akers, Landen had an average of 14 more receiving yards
from the previous year.

As for defense we can see that Rose, Mike had an average of 1.5 more solo tackles in 2020.
