---
title: "team.Rmd"
author: "Dana Thacker, Andrew Fahmy"
date: "3/15/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(readxl)
library(tidyverse)
library(ggplot2)
defense <- read_excel('../ds202_lab4/cyclonesFootball2020.xlsx', sheet='Defensive')
offense <- read_excel('../ds202_lab4/cyclonesFootball2020.xlsx', sheet='Offensive')
biography <- read_excel('../ds202_lab4/cyclonesFootball2020.xlsx', sheet='Biography')
```

**Part One: Cleaning Data**
Although the three Excel sheets are loaded into R, they are not useful before some cleaning. Manipulate the three imported data frames, so that in the cleaned dataset:
*Problem 1*
The player names and opponent names are factors.
```{r}
defense <- defense %>%
  mutate(Name = as.factor(Name), 
         Opponent_Opponent = as.factor(Opponent_Opponent))

offense <- offense %>%
  mutate(Name = as.factor(Name), 
         Opponent_Opponent = as.factor(Opponent_Opponent))

biography <- biography %>%
  mutate(Name = as.factor(Name))
```

*Problem 2*
All offensive and defensive statistics are numerical. The Weight column in biography is also numeric. (Optional: Read ?across and learn to mutate multiple columns simultaneously using the across function in dplyr)
```{r}
defense <- defense %>%
  mutate(across(Tackles_Solo:Pass_PB, as.numeric))

offense <- offense %>%
  mutate(across(Receiving_REC:Passing_INT, as.numeric))

biography <- biography %>%
  mutate(Weight = as.numeric(Weight))
```

*Problem 3*
Change the Height column in biography into numeric. (Hint: Use a unit that works well)
```{r}
biography <- biography %>%
  separate(Height, c('Feet', 'Inches'), sep='-') 

biography <- biography %>%
  mutate(Feet = as.numeric(Feet),
         Inches = as.numeric(Inches))
```

**Part 2: Tidying Data**
*Problem 1*
Reformat the offClean data frame into a tidy format using pivot_longer. The type of statistic (Receiving_REC for the number of receptions, Receiving_YDS for receiving yards, etc) is added as a new key column named stat.
```{r}
offClean <- offense %>%
  pivot_longer(Receiving_REC:Passing_INT, names_to = 'Type', values_to = 'Stat')
```

*Problem 2*
Show on the distributions of the offensive statistics. The distribution of each offensive statistics should be shown in a separate facet. Explain the patterns you see.
```{r}
ggplot(offClean, aes(x = Stat)) + geom_histogram() + facet_wrap(~ Type, scales = 'free') + xlab("Stat") + ylab("Count") + ggtitle("Statistics by Type")
```

*Problem 3*
Did ISU have better offense against Oregon or Oklahoma? Answer this question by creating a scatterplot with the x- and y-axes being the receiving yards of the ISU players against these two opponents. A large number of receiving yards indicates good offense.
```{r}
offOreOkl <- offClean %>%
  filter(Opponent_Opponent %in% c("Oregon", "Oklahoma")) %>%
  group_by(Name) %>%
  filter(Type == "Receiving_YDS")

offOklahoma <- offOreOkl %>%
  filter(Opponent_Opponent %in% c("Oklahoma")) %>%
  group_by(Name) %>%
  summarise(Average = mean(Stat))

offOregon <- offOreOkl %>%
  filter(Opponent_Opponent %in% c("Oregon"))

offOreOklAvg <- offOreOkl %>%
  left_join(offOklahoma)

offOreOklPivoted <- offOreOklAvg %>%
  pivot_wider(names_from = 'Type', values_from = 'Stat') %>%
  filter(Opponent_Opponent %in% c("Oregon")) %>%
  rename(Oklahoma_Receiving_YDS = Average,
         Oregon_Receiving_YDS = Receiving_YDS) %>%
  mutate(Oklahoma_Receiving_YDS = as.numeric(Oklahoma_Receiving_YDS),
         Oregon_Receiving_YDS = as.numeric(Oregon_Receiving_YDS))
  
ggplot(offOreOklPivoted, aes(x = Oklahoma_Receiving_YDS, y = Oregon_Receiving_YDS)) + geom_point() + xlab("Average Oklahoma Receiving Yards") + ylab("Oregon Receiving Yards") + ggtitle("Oklahoma Receiving Yards vs Oregon Receiving Yards by ISU in 2020")
```

*Problem 4*
Separate the Hometown column into two columns, namely, the city and the state where the player comes from. Print the first few rows of these two columns. (Hint: look at the sep= argument of separate)
```{r}
biography <- biography %>%
  separate(Hometown, c('City', 'State'), sep = ', ')

head(biography, 4)
```

*Problem 5*
How many players are there from each state?
```{r}
ggplot(biography, aes(x = State)) + geom_bar() + xlab("State") + ylab("Count") + ggtitle("Number of Players From Each State")
```

**Part 3: Joining Data Frames**
*Problem 1*
Find which states do the ISU players (both defensive and offensive) come from. Create an appropriate numerical or graphical summary.
```{r}
ggplot(biography, aes(x = State)) + geom_bar() + xlab("State") + ylab("Count") + ggtitle("Number of Players From Each State")
```

*Problem 2*
How does the performance of quarterback Brock Purdy relate to the performance of the ISU team?
```{r}


```

*Problem 3*
Which ISU player(s) made the biggest progress over the last year? Compare the player-level statistics (i.e. averages and totals for each player) using both the 2020 and the 2019 data here. Choose for yourself the player statistics to investigate.
```{r}



```

